GPU-friendly Stroke Expansion

### 2024-04-29 paper.tex
Add a actual vs estimated memory table (which I forgot to do earlier) and expand discussion

\begin{table}
    \caption{Comparison of input and output segment counts in the more intensive GPU test cases. \emph{Segments} are input path segments. \emph{Lines} and \emph{Arcs} are the output primitives.}
    \begin{tabular}{|l|c|c|c|}
    \hline
        \multicolumn{1}{|c|}{\textbf{Test}} & \textbf{Input Segments} & \textbf{Lines} & \textbf{Arcs} \\
    \hline
        waves.svg & 13,308 & 406,059 & 111,965 \\
        mmark-70k & 70,000 & 1,039,899 & 500,351 \\
        mmark-120k & 120,000 & 1,784,896 & 858,162 \\
        long path (butt) & 503,304 & 1,060,476 & 1,060,476 \\
        long path (round) & 503,304 & 2,013,216 & 1,060,476 \\
    \hline
    \end{tabular}
    \label{table:bump-counts}
\end{table}

@@

Figure~\ref{fig:tag-monoid-gpu-timings} shows the run time of the tag monoid parallel prefix sums with increasing input size. The prefix sums scale better at large input sizes compared to our stroke expansion kernel. This can be attributed to a lack of expensive computations and high control-flow uniformity in the tag monoid kernel.

@@

We augmented our encoding logic to compute a conservative estimate of the required output buffer size based on an upper bound on the number of line primitives per input path segment. We found that Wang's formula works well in practice, as it is relatively inexpensive to evaluate, though it results in a higher memory estimate than required (see Table~\ref{table:bump-estimates}). For parallel curves, Wang's theorem does not guarantee an accurate upper bound on the required subdivisions, which we worked around by inflating the estimate for the source by a scale factor derived from the linewidth. We observed that enabling estimation increases our CPU pre-processing time by 2-3x on the largest scenes but the impact is overall negligible when the scene size is modest.

\begin{figure}
    \includegraphics[scale=0.66]{tag_monoid_gpu_timings}
    \caption{Mean GPU execution time of the tag monoid parallel prefix sums with increasing number of input path segments, measured on Apple M1 Max and NVIDIA RTX 4090. The figure also shows the execution time of the stroke expansion and flattening kernel and compares their scaling with input size. The y-axis is shown in logarithmic scale to highlight the similar scaling trend on both GPUs despite the order-of-magnitude difference in performance.}
    \Description{Mean GPU execution time of the tag monoid kernel}
    \label{fig:tag-monoid-gpu-timings}
\end{figure}

\begin{table}
    \caption{Comparison of CPU encoding time with and without size estimation on the largest scenes. All measurements were computed on an Apple M1 Max. The additional cost is negligible when the scene size is modest but tangible with increased complexity.}
    \begin{tabular}{|l|c|c|}
    \hline
        \multicolumn{1}{|c|}{\textbf{Test}} & \textbf{Encoding} & \textbf{Encoding + Estimation} \\
    \hline
        waves.svg  & 6 $\mu$s & 6.13 $\mu$s \\
        mmark-70k  & 2.32 ms  & 6.59 ms     \\
        mmark-120k & 4.17 ms  & 11.26 ms    \\
        long path  & 4.14 ms  & 6.28 ms     \\
    \hline
    \end{tabular}
    \label{table:bump-estimates}
\end{table}

@@

    \item The pre-allocation requirement for bump-allocated line soup buffer is a limitation due to today's graphics APIs. An interesting future exploration is to look at newer GPU API features, such as mesh shaders and work graphs, for GPU-driven scheduling of the compute workload under a bounded memory footprint.

>>>>

\begin{table}
    \caption{Comparison of input and output segment counts in the more intensive GPU test cases. \emph{Segments} are input path segments. \emph{Lines} and \emph{Arcs} are the output primitives.}
    \begin{tabular}{|l|c|c|c|}
    \hline
        \multicolumn{1}{|c|}{\textbf{Test}} & \textbf{Input Segments} & \textbf{Lines} & \textbf{Arcs} \\
    \hline
        waves.svg & 13,308 & 406,059 & 111,965 \\
        mmark-70k & 70,000 & 1,039,899 & 500,351 \\
        mmark-120k & 120,000 & 1,784,896 & 858,162 \\
        long dash (butt) & 503,304 & 1,060,476 & 1,060,476 \\
        long dash (round) & 503,304 & 2,013,216 & 1,060,476 \\
    \hline
    \end{tabular}
    \label{table:bump-counts}
\end{table}

@@

Figure~\ref{fig:tag-monoid-gpu-timings} shows the run time of the tag monoid parallel prefix sums with increasing input size. The prefix sums scale better at large input sizes compared to our stroke expansion kernel. This can be attributed to a lack of expensive computations and high control-flow uniformity in the tag monoid kernel. The CPU encoding times (prior to the GPU prefix sums) for some of the large scenes is shown in Table~\ref{table:bump-estimate-cost}. Dashing on the GPU was out of scope, so the dash style applied to the \emph{long dash} scene had to be processed on the CPU and dashes were sequantially encoded as individual strokes. Including CPU-side dashing, the time to encode \emph{long dash} on the Apple M1 Max was approximately 25 ms on average, while the average time to encode the path after dashing was 4.14 ms.

@@

We augmented our encoding logic to compute a conservative estimate of the required output buffer size based on an upper bound on the number of line primitives per input path segment. We found that Wang's formula works well in practice, as it is relatively inexpensive to evaluate, though it results in a higher memory estimate than required (see Table~\ref{table:bump-estimate-mem}). For parallel curves, Wang's theorem does not guarantee an accurate upper bound on the required subdivisions, which we worked around by inflating the estimate for the source by a scale factor derived from the linewidth. We observed that enabling estimation increases our CPU pre-processing time by 2-3x on the largest scenes but the impact is overall negligible when the scene size is modest.

\begin{figure}
    \includegraphics[scale=0.66]{tag_monoid_gpu_timings}
    \caption{Mean GPU execution time of the tag monoid parallel prefix sums with increasing number of input path segments, measured on Apple M1 Max and NVIDIA RTX 4090. The figure also shows the execution time of the stroke expansion and flattening kernel. The y-axis is shown in logarithmic scale to highlight the similar scaling trend on both GPUs despite the order-of-magnitude difference in performance.}
    \Description{Mean GPU execution time of the tag monoid kernel}
    \label{fig:tag-monoid-gpu-timings}
\end{figure}

\begin{table}
    \caption{Comparison of CPU encoding time with and without size estimation on the largest scenes. All measurements were computed on an Apple M1 Max. The additional cost is negligible when the scene size is modest but tangible with increased complexity.}
    \begin{tabular}{|l|c|c|}
    \hline
        \multicolumn{1}{|c|}{\textbf{Test}} & \textbf{Encoding} & \textbf{Encoding + Estimation} \\
    \hline
        roads.svg  & 398.56 $\mu$s & 618.24 $\mu$s \\
        waves.svg  & 197.53 $\mu$s & 418.96 $\mu$s \\
        mmark-70k  & 2.69 ms       & 6.89 ms       \\
        mmark-120k & 4.28 ms       & 11.59 ms      \\
    \hline
    \end{tabular}
    \label{table:bump-estimate-cost}
\end{table}

\begin{table}
    \caption{The estimated and actual \emph{line soup} buffer sizes for scenes rendered at 2088x1600 resolution. Our estimation uses Wang's formula combined with a scaling heuristic based on the linewidth. In the worst case, our buffer size estimator overshoots by a factor ~4.}
    \begin{tabular}{|l|c|c|}
    \hline
        \multicolumn{1}{|c|}{\textbf{Test}} & Actual Size & Estimated Size \\
    \hline
        roads.svg  & 958.78 KB & 2.45 MB  \\
        waves.svg  & 9.37 MB   & 10.44 MB \\
        mmark-70k  & 23.81 MB  & 100.79   \\
        mmark-120k & 40.86 MB  & 173.56   \\
    \hline
    \end{tabular}
    \label{table:bump-estimate-mem}
\end{table}

@@

    \item The pre-allocation requirement for bump-allocated line soup buffer is a limitation due to today's graphics APIs. A more accurate and optimized buffer size estimation heuristic is considered future work. Another interesting future exploration is to employ newer GPU API features, such as mesh shaders and work graphs, for GPU-driven scheduling of the compute workload under a bounded memory footprint.
